---
output:
  pdf_document:
    latex_engine: xelatex
header-includes:
  - \usepackage{xcolor}
  - \usepackage{graphicx}
  - \usepackage{tcolorbox} % For creating text boxes
  - \usepackage{xcolor} % Use color package for customization
  - \usepackage{varwidth}
  - \definecolor{whodarkblue}{RGB}{0, 32, 92}
  - \definecolor{wholightblue}{RGB}{0, 154, 222}
  - \definecolor{wholightblueshade}{RGB}{121, 181, 227}
  - \definecolor{wholightblueshadev2}{RGB}{227, 235, 244}
  - \setmainfont{Arial}
  - \usepackage{fancyhdr} % Allows customization of headers and footers
  - \pagestyle{fancy} % Enable fancy header/footer style
  - \fancyhf{} # Clear all default header/footer
  # - \fancyhead{} % Clear default header
  - \fancyhead[L]{\color{whodarkblue}Country Health Profile}
  - \fancyhead[C]{} % Empty center
  - \fancyhead[R]{\color{whodarkblue}\thepage} % Add right-aligned content (e.g., author name)
  - \renewcommand{\headrule}{{\color{whodarkblue}\hrule width\headwidth height\headrulewidth}}
  - \renewcommand{\headrulewidth}{0.6pt} % Add a horizontal line with thickness of 0.4pt
  - \setlength{\headheight}{30pt} % Increase header height if needed
geometry: margin=1.75cm
params:
  country: NA
  country_iso: NA
  filtered_indicator_values: NA
  country_table: NA
  all_plt_dat: NA
  inds_all: NA
  ind_labels: NA
  config_all: NA
  country_ref: NA
---
```{r parameters, echo = FALSE, message = FALSE, include = FALSE, eval = TRUE}
list2env(params, .GlobalEnv)
```
\thispagestyle{empty}
\pagecolor{whodarkblue}

\begin{flushleft}
{\color{white}\rule{\linewidth}{0.25mm}} \\
\vspace{1.5cm}
{\fontsize{40}{50}\selectfont \color{white} Country Health Profile} \\
\vspace{1.5cm}
{\fontsize{25}{30}\selectfont \color{wholightblueshade} `r toupper(country)`}

\vspace*{15cm} 

\end{flushleft}

\begin{flushright}
\includegraphics[width=5cm]{WHO-EN-W-H.png}
\end{flushright}

<!-- \begin{flushright} -->
<!-- {\fontsize{9}{10}\selectfont \color{white} Last updated 10 February 2025}\\ -->
<!-- \end{flushright} -->


\newpage
\pagecolor{white}

```{r settings, echo= FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(treemap)
library(ghost)
library(sf)
library(RColorBrewer)
library(kableExtra)
library(ggpol)
library(extrafont)
library(showtext)
library(patchwork)
library(cowplot)
library(ggnewscale)
library(flextable)
library(whoville)
library(DemoDecomp)
library(ggrepel)
library(ggbump)
library(forcats)
library(ggpubr)
library(ggh4x)
library(httr)
library(jsonlite)


inds <- get_inds(get_forecast_inds = FALSE, get_billion_inds = TRUE)
inds$uhc <- c(inds$uhc, "population")
inds$hep <- config_all$hep_intermediate_inds
inds_summary <- c("hpop" = "hpop_healthier", "uhc" = "uhc_billion", "hep" = "hep_idx")

inds_all <- inds %>%
  unlist(use.names = FALSE) %>%
  unique() %>%
  c(inds_summary) %>%
  unname()

# country = "Thailand"; country_iso = "THA"

`%!in%` = Negate(`%in%`)
font_add("calibri", regular = "fonts/calibri/calibri.ttf")
showtext_auto()

theme_set(theme(text = element_text(family="calibri"),
                title = element_text(family="calibri"),
                axis.text.x = element_text(family = "calibri"),
                legend.text = element_text(family="calibri"),
          axis.text.y = element_text(family = "calibri"),
          axis.title.x = element_text(family = "calibri"),
          axis.title.y = element_text(family = "calibri")))

```

```{r region_info, include=FALSE}
who_region <- country_table %>% filter(Code == country_iso) %>% pull(ParentCode)
# who_region = "AMR"
who_region_name <- case_when(
  who_region=="AFR" ~ "African Region",
  who_region=="AMR" ~ "Region of the Americas",
  who_region=="EMR" ~ "Eastern Mediterranean Region",
  who_region=="EUR" ~ "European Region",
  who_region=="SEAR" ~ "South-East Asia Region",
  who_region=="WPR" ~ "Western Pacific Region")
who_region_countries <- country_table %>% filter(ParentCode == who_region) %>% pull(Code) %>% substr(1,3) %>% unique()
```

```{r data, include=FALSE}
year <- 2021
comp_year <- 2000
country_ <- country 


response <- GET("https://xmart-api-public-uat.who.int/REFMART/REF_COUNTRY")
resultTxt <- content(response, "text")
resultJson <- fromJSON(resultTxt, flatten = T)
country_ref <- resultJson$value %>%
  mutate(m49code = case_when(
    CODE_ISO_NUMERIC<10 ~ paste0("00", CODE_ISO_NUMERIC),
    CODE_ISO_NUMERIC>=10 & CODE_ISO_NUMERIC<=99 ~ paste0("0", CODE_ISO_NUMERIC),
    TRUE ~ as.character(CODE_ISO_NUMERIC)
  ))

m49code <- country_ref %>% filter(CODE_ISO_3==country_iso) %>% pull(m49code)


getdata <- function(ind){
  x <- GET(url = paste("https://ghoapi.azureedge.net/api/", ind, sep=""))
  x2 = fromJSON(rawToChar(x$content))
  x3 <- x2[2]$value

  return(x3)
}

le <- getdata("WHOSIS_000001")
hale <- getdata("WHOSIS_000002")

lehale <- rbind(le, hale) %>%
  mutate(sex = case_when(
    Dim1=="SEX_BTSX" ~ "Both sexes",
    Dim1=="SEX_MLE" ~ "Male",
    Dim1=="SEX_FMLE" ~ "Female",
  )) %>%
  select(-c(TimeDimensionBegin, TimeDimensionEnd, Date, Comments, Dim1, Dim1Type, Dim2Type, Dim2, Dim3Type, Dim3, DataSourceDimType, DataSourceDim, Id)) %>%
  rename(iso3 = SpatialDim, year = TimeDim, value = NumericValue) %>%
  mutate(country = whoville::iso3_to_names(iso3)) %>% 
  filter(iso3 == country_iso)


pop_age <- read.csv("population_estimates_byage.csv")


url <- paste0("https://xmart-api-public.who.int/DEX_CMS/GHE_FULL?$filter=DIM_COUNTRY_CODE%20eq%20'", country_iso, "'%20and%20DIM_SEX_CODE%20eq%20'TOTAL'%20and%20FLAG_RANKABLE%20eq%201") #%20and%20FLAG_TREEMAP%20eq%201
response <- GET(url)
resultTxt <- content(response, "text")
resultJson <- fromJSON(resultTxt, flatten = T)
rwa <- resultJson$value
```

```{r text_values, include=FALSE}
pop <- read.csv("population_estimates.csv")

# total population
total_pop2023 <- as.numeric(pop %>% filter(m49code_ == as.numeric(m49code) & year==2023) %>% pull(totalpop_jul))
total_pop2050 <- as.numeric(pop %>% filter(m49code_ == as.numeric(m49code) & year==2050) %>% pull(totalpop_jul))
total_pop_increase <- abs(round(((as.numeric(total_pop2050) - as.numeric(total_pop2023)) / as.numeric(total_pop2023))*100, digits = 0))
popchange_text <- case_when(
 total_pop2050 < total_pop2023 ~ "decrease",
 total_pop2023 <= total_pop2050 ~ "increase"
)

life_exp <- lehale %>% filter(IndicatorCode=="WHOSIS_000001" & sex=="Both sexes" & year==2021 & iso3==country_iso) %>% pull(value)
life_exp_2000 <- lehale %>% filter(IndicatorCode=="WHOSIS_000001" & sex=="Both sexes" & year==2000 & iso3==country_iso) %>% pull(value)

# total number of deaths
total_deaths <- cod19 %>%
    filter(sex=="Both sexes" & DIM_AGEGROUP_CODE == "TOTAL" & DIM_COUNTRY_CODE == country_iso & FLAG_LEVEL == 0) %>%
    pull(VAL_DEATHS_COUNT_NUMERIC)

# top causes of death
top_death_3 <- cod19 %>%
  filter(sex=="Both sexes" & DIM_AGEGROUP_CODE == "TOTAL" & DIM_COUNTRY_CODE == country_iso & FLAG_TREEMAP==1) %>%
  arrange(desc(VAL_DEATHS_COUNT_NUMERIC)) %>% 
  filter(row_number()<=3) %>% 
  pull(DIM_GHECAUSE_TITLE)

```

\begin{tcolorbox}[colback=wholightblueshadev2, colframe=wholightblueshadev2, title=Overview, coltitle=whodarkblue, fonttitle=\large\bfseries, height=4.5cm, boxsep=2mm]
\begin{varwidth}{\linewidth}
In `r country_`, the population as of 2023 is `r format(total_pop2023*1000, big.mark = " ")` with a projected `r popchange_text` of `r total_pop_increase`\% to `r format(total_pop2050*1000, big.mark = " ")` by 2050. 
\end{varwidth}
\newline \newline
\begin{varwidth}{\linewidth}
The life expectancy at birth was `r round(life_exp, 1)` years in 2021, an increase from `r round(life_exp_2000, 1)` in 2000.
\end{varwidth}
\newline \newline
\begin{varwidth}{\linewidth}
`r country` had `r format(round(total_deaths), big.mark=" ", scientific=F)` total deaths in 2021. The top causes of death were: `r top_death_3[1]`, `r top_death_3[2]`, and, `r top_death_3[3]`.
\end{varwidth}
\end{tcolorbox}
<!-- \rule{\textwidth}{0.6pt}\color{whodarkblue} -->
\color{whodarkblue}
```{r pop le, fig.width=7.5, fig.align='center', fig.height=3.5}
pop_data <- pop_age %>% 
  filter(country == country_ & year %in% c(2023, 2050)) %>% 
  mutate(ageg = factor(ageg, 
                       levels = c("0-4", "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", "35-39",
                                  "40-44", "45-49", "50-54", "55-59", "60-64", "65-69", "70-74", "75-79", "80-84", "85+")
  ))

pop_data2 <- pop_data %>%
  mutate(age = as.numeric(str_extract(ageg, "^[:digit:]+"))) %>% 
  mutate(ageg2 = case_when(
    age == 0 ~ "0-4",
    age >4 & age <15 ~ "5-14",
    age >=15 & age <45 ~ "15-44",
    age >=45 & age <60 ~ "45-59",
    age>=60 ~ "60+"
  )) %>% 
  mutate(ageg2 = factor(ageg2, levels = c("0-4","5-14","15-44","45-59","60+"))) %>% 
  group_by(country, Location.code, year, ageg2) %>% 
  summarise(pop = sum(total_pop)) %>% 
  ungroup() %>% 
  group_by(country, Location.code, year) %>% 
  mutate(prop = round((pop/sum(pop))*100,0),
         total_pop = round(sum(pop)*1000, 0)) %>% 
  ungroup() 
  
max_lim <- plyr::round_any(max(pop_data2 %>% filter(year==2023) %>% pull(pop)), 500, f = ceiling)

pop <- ggplot() + 
  geom_col(data = pop_data2 %>% filter(year==2023),
           aes(x= ageg2, y = pop), width = 0.4, show.legend = F, fill = "#009ADE") +
  coord_flip()+
  geom_text(data = pop_data2 %>% filter(year==2023),
            aes(x= ageg2, y = pop, label = paste0(prop, "%")), color = "white", size = 2.5, hjust =1)+
  labs(subtitle = "Population by age (thousands), 2023")+
  scale_y_continuous(expand = c(0,0), limits = c(0, max_lim))+
  theme_classic() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_line(color = "white"),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    panel.grid.major.x = element_line(color = "grey"),
    axis.text.y = element_text(family="calibri", size = 8, color = "black"),
    plot.subtitle = element_text(size = 9, color = "#00205C"),
    plot.margin = margin(t = 0, l = 0, r = 20, b = 0, unit = "pt"),
  )  

exp <- lehale %>% 
  filter(IndicatorCode=="WHOSIS_000002" | IndicatorCode=="WHOSIS_000001") %>%
  filter(year %in% c(2000, 2010, 2019, 2020, 2021)) %>%
  # filter(country %in% c(country_, who_region_name, "Global")) %>%
  filter(iso3 == country_iso) %>%
  filter(sex %in% c("Both sexes")) %>% 
  mutate(value = round(value, 1)) %>% 
  mutate(name2 = case_when(
    IndicatorCode=="WHOSIS_000001" ~ "Life expectancy",
    IndicatorCode=="WHOSIS_000002" ~ "Healthy life expectancy",
  )) %>% 
  mutate(name2 = factor(name2, levels = c("Life expectancy", "Healthy life expectancy"))) %>% 
  select(country, year, sex, name2, value) 

exp2 <- exp %>% pivot_wider(names_from = name2, values_from = value)

max <- exp %>% select(value) %>% arrange(desc(value)) %>% filter(row_number()==1) %>% pull()
min <- exp %>% select(value) %>% arrange(value) %>% filter(row_number()==1) %>% pull()

lehale_g <- exp %>% 
  ggplot(aes(x=year, y=value, color=name2))+
  geom_line(size = 1, show.legend=F) +
  geom_point(size = 2, show.legend=F) +
  geom_text_repel(data = exp %>% filter(year==2010 & country == country_),
                  aes(x=year, y=value, color=name2, label = str_wrap(name2, 15)),
                  na.rm = T,
                  direction = "y",
                  nudge_y = 1,
                  # xlim = c(2021, Inf),
                  max.overlaps = 100,
                  show.legend = FALSE,
                  size = 2.5)+
  labs(subtitle = "Life expectancy and healthy life expectancy, 2000-2021") +
  scale_y_continuous(limits = c(round(min, 0)-2,round(max, 0)+6)) +
  scale_x_continuous(breaks = c(2000, 2010, 2020), limits = c(2000, 2021))+
  scale_color_manual(values = c("Life expectancy" ="#009ADE",
                                "Healthy life expectancy" = "#80BC00")) +
  theme_classic()+
  theme(axis.text=element_text(size=8),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5, color = "black"),
        axis.text.y = element_text(size=8, color = "black"),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        # plot.margin = margin(t = 0, l = 0, r = 10, b = 0, unit = "pt"),
        plot.subtitle = element_text(size = 9, color = "#00205C"),
        panel.border = element_blank(),
        axis.line.y = element_blank(),
        axis.line.x = element_line(color = "darkgrey",linetype = 3),
        panel.grid.major = element_line(color = "darkgrey",size = 0.5,linetype = 3),
        panel.grid = element_blank(),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        legend.title = element_blank()
        )

wrap_plots(pop, lehale_g, ncol = 2)
```

\normalsize Top 10 causes of death, 2021
```{r cod, fig.height=3.6, fig.align='center'}

url2 <- paste0("https://xmart-api-public.who.int/DEX_CMS/GHE_FULL?$filter=DIM_COUNTRY_CODE%20eq%20'", country_iso, "'%20and%20DIM_SEX_CODE%20eq%20'TOTAL'%20and%20FLAG_LEVEL%20eq%200%20and%20DIM_YEAR_CODE%20eq%202021") 
response <- GET(url2)
resultTxt <- content(response, "text")
resultJson <- fromJSON(resultTxt, flatten = T)

total_deaths_u70 <- resultJson$value %>%
  filter(DIM_AGEGROUP_CODE %!in% c("Y70T74", "Y75T79", "Y80T84", "YGE_85", "TOTAL", "D0T27", "M1T11")) %>% 
  group_by(DIM_COUNTRY_CODE, DIM_YEAR_CODE, DIM_GHECAUSE_TITLE) %>% 
  summarise(dth = round(sum(VAL_DTHS_COUNT_NUMERIC),0)) %>% 
  pull(round(dth,0)) 

#-------
top10_ <- rwa %>%
  filter(DIM_AGEGROUP_CODE=="TOTAL" & DIM_YEAR_CODE==2021) %>% 
  group_by(DIM_COUNTRY_CODE, DIM_GHECAUSE_TITLE) %>% 
  summarise(deaths = sum(VAL_DTHS_COUNT_NUMERIC)) %>% 
  ungroup() %>% 
  group_by(DIM_COUNTRY_CODE) %>%
  arrange(desc(deaths)) %>%
  slice(1:10) %>% 
  pull(DIM_GHECAUSE_TITLE)

rwa_cod_un70 <- rwa %>%
  filter(DIM_YEAR_CODE==2021 &
           DIM_AGEGROUP_CODE %!in% c("Y70T74", "Y75T79", "Y80T84", "YGE_85", "TOTAL", "D0T27", "M1T11") & 
           DIM_GHECAUSE_TITLE %in% top10_) %>% 
  group_by(DIM_COUNTRY_CODE, DIM_YEAR_CODE, DIM_GHECAUSE_TITLE) %>% 
  summarise(dths_u70 = round(sum(VAL_DTHS_COUNT_NUMERIC),0),
            pop_u70 = sum(ATTR_POPULATION_NUMERIC)) %>%
  ungroup() %>%
  mutate(dth_rate_u70 = round((dths_u70/pop_u70)*100000, 1),
         dth_perc_u70 = round((dths_u70/total_deaths_u70)*100,0))

rwa_cod <- rwa %>% 
  filter(DIM_YEAR_CODE==2021 &
           DIM_AGEGROUP_CODE == "TOTAL" & 
           DIM_GHECAUSE_TITLE %in% top10_) %>% 
  arrange(desc(VAL_DTHS_COUNT_NUMERIC)) %>% 
  select(DIM_COUNTRY_CODE, DIM_YEAR_CODE, DIM_GHECAUSE_TITLE, 
         dths_all = VAL_DTHS_COUNT_NUMERIC, 
         dth_rate_all = VAL_DTHS_RATE100K_NUMERIC, 
         pop_all = ATTR_POPULATION_NUMERIC) %>% 
  mutate(dths_all = round(dths_all,0),
         dth_rate_all = round(dth_rate_all,1),
         dth_perc_all = round((dths_all/total_deaths)*100,0),
         Country = "Rwanda") %>% 
  left_join(rwa_cod_un70, by = c("DIM_COUNTRY_CODE", "DIM_YEAR_CODE", "DIM_GHECAUSE_TITLE")) 

total_pop_u70 <- rwa_cod %>% pull(pop_all) %>% unique()

total_pop <- rwa_cod %>% pull(pop_u70) %>% unique()


rwa_cod2 <- rwa_cod %>% 
  select(-c(DIM_COUNTRY_CODE, DIM_YEAR_CODE, pop_all, pop_u70, Country)) %>% 
  rbind(
    data.frame(
      DIM_GHECAUSE_TITLE = "All causes, total",
      dths_all = round(total_deaths,0),
      dth_rate_all = round((total_deaths/total_pop)*100000, 1),
      dth_perc_all = 100,
      dths_u70 = total_deaths_u70,
      dth_rate_u70 = round((total_deaths_u70/total_pop_u70)*100000, 1),
      dth_perc_u70 = 100
    )
  ) %>% 
  select(`Cause of death` = DIM_GHECAUSE_TITLE,
         `Number of deaths, total` = dths_all,
         `Rate per 100,000, total` = dth_rate_all,
         # `Percent (%), total` = dth_perc_all,
         `Number of deaths` = dths_u70,
         `Rate per 100,000` = dth_rate_u70
         # `Percent (%)` = dth_perc_u70
)

library(officer)

flextable(rwa_cod2) %>% 
  # theme_zebra(odd_body = "#eef3f7", odd_header = "#eef0f1") %>%
  bg(., i = NULL, j = NULL, "#e6e7e9", part = "header") %>% # "#e3ebf4",
  border_remove(.) %>% 
  set_header_labels(., 
                    `Number of deaths, total` = "Number of deaths",
                    # `Percent (%), total` = "Percent (%)",
                    `Rate per 100,000, total` = "Rate per 100,000"
  ) %>% 
  add_header_row(top = TRUE, values = c("", "Total", "Under 70 years old"), colwidths = c(1, 2, 2)) %>% 
  width(., j =1, width=6, unit = "cm") %>%
  width(., j =2:5, width=3, unit = "cm") %>%
  align(., j=2:5, align = "center") %>% 
  align(., j=2:5, align = "center", part = "header") %>% 
  bold(., bold = TRUE, part = "body", i = 11) %>% 
  hline_bottom(., border = fp_border(color = "grey", width = 1), part = "body") %>%
  hline_top(., border = fp_border(color = "grey", width = 1), part = "all") %>%
  # hline_top(., border = fp_border(color = "black", width = 1), part = "header") %>%
  fontsize(., size = 9, part = "all")
```
\tiny Data source: UNDESA World Population Prospects 2022 & WHO Global Health Estimates 2024
\newpage
\normalsize Change in the top 10 broad causes of death, 2000-2021
``` {r top10, fig.height=3.7, fig.align='center'}

url <- paste0("https://xmart-api-public.who.int/DEX_CMS/GHE_FULL?$filter=DIM_COUNTRY_CODE%20eq%20'", country_iso, "'%20and%20DIM_SEX_CODE%20eq%20'TOTAL'%20and%20FLAG_LEVEL%20eq%202%20and%20DIM_AGEGROUP_CODE%20eq%20'TOTAL'")
response <- GET(url)
resultTxt <- content(response, "text")
resultJson <- fromJSON(resultTxt, flatten = T)
cod_ <- resultJson$value %>% filter(DIM_YEAR_CODE %in% c(2000, 2021))


b <- cod_ %>% 
  filter(DIM_YEAR_CODE==2021) %>% 
  select(DIM_GHECAUSE_TITLE, VAL_DTHS_COUNT_NUMERIC, DIM_YEAR_CODE) %>% 
  mutate(prop_deaths = VAL_DTHS_COUNT_NUMERIC / sum(VAL_DTHS_COUNT_NUMERIC),
         rank_deaths = rank(-prop_deaths, ties.method = "first")) %>% 
  rbind(cod_ %>% 
          filter(DIM_YEAR_CODE==2000) %>% 
          select(DIM_GHECAUSE_TITLE, VAL_DTHS_COUNT_NUMERIC, DIM_YEAR_CODE) %>% 
          mutate(prop_deaths = VAL_DTHS_COUNT_NUMERIC / sum(VAL_DTHS_COUNT_NUMERIC),
                 rank_deaths = rank(-prop_deaths, ties.method = "first"))) %>% 
  pivot_wider(id_cols = DIM_GHECAUSE_TITLE, names_from = DIM_YEAR_CODE,
                  values_from = c(rank_deaths)) %>%
  filter(`2000`<11 | `2021`<11) %>% 
  pivot_longer(cols = c(`2000`, `2021`)) %>% 
  mutate(year = as.numeric(name),
         label = paste0(value, ". ", DIM_GHECAUSE_TITLE))
    
b %>% 
  ggplot(aes(x = year, y = -value, group = DIM_GHECAUSE_TITLE)) + #, color = rank_change
  geom_bump(size = 3, show.legend = F, alpha = 0.4, color = "#009ADE") +
  geom_bump(size = 0.6, show.legend = F, alpha = 1, color = "#009ADE") +
  geom_point(size = 6, show.legend = F, color = "#009ADE") +
  geom_text(data = b %>% filter(year == 2000),
            aes(x = year, y = -value, label = value),
            size = 3, show.legend = F, color = "black") +
  geom_text(data = b %>% filter(year == 2021),
            aes(x = year, y = -value, label = value),
            size = 3, show.legend = F, color = "black") +
  geom_text(data = b %>% filter(year == 2000),
            aes(x = year - 1, y = -value, label = str_wrap(DIM_GHECAUSE_TITLE,25)),
            size = 3, hjust = 1, show.legend = F, color = "black", lineheight = 0.85) +
  geom_text(data = b %>% filter(year == 2021),
            aes(x = year + 1, y = -value, label = str_wrap(DIM_GHECAUSE_TITLE,25)),
            size = 3, hjust = 0, show.legend = F, color = "black", lineheight = 0.85) +
  scale_x_continuous(limits = c(1985, 2036))+
  theme_void() +
  guides(color=guide_legend(ncol=1))
```
\
\rule{\textwidth}{0.6pt}\color{whodarkblue}
\
\normalsize Broad causes of years of life lost (YLL), 2021
``` {r top10_yll, fig.height=4, fig.align='center'}

url <- paste0("https://xmart-api-public.who.int/DEX_CMS/GHE_FULL?$filter=DIM_COUNTRY_CODE%20eq%20'", country_iso, "'%20and%20DIM_SEX_CODE%20eq%20'TOTAL'%20and%20DIM_AGEGROUP_CODE%20eq%20'TOTAL'%20and%20FLAG_LEVEL%20eq%202")
response <- GET(url)
resultTxt <- content(response, "text")
resultJson <- fromJSON(resultTxt, flatten = T)
yll <- resultJson$value %>% rename(FLAG_CAUSEGROUP = DIM_CAUSE_GROUP)

yll_ <- yll %>% 
  filter(DIM_COUNTRY_CODE == country_iso & FLAG_LEVEL==2 & DIM_YEAR_CODE==2021) %>%
  select(DIM_GHECAUSE_TITLE, VAL_YLL_COUNT_NUMERIC, DIM_YEAR_CODE, FLAG_CAUSEGROUP) %>% 
  mutate(prop_yll = VAL_YLL_COUNT_NUMERIC / sum(VAL_YLL_COUNT_NUMERIC),
         rank_yll = rank(-prop_yll, ties.method = "first")) %>% 
  mutate(yll = round(VAL_YLL_COUNT_NUMERIC, 0))


url <- paste0("https://xmart-api-public.who.int/DEX_CMS/GHE_FULL?$filter=DIM_COUNTRY_CODE%20eq%20'", country_iso, "'%20and%20DIM_SEX_CODE%20eq%20'TOTAL'%20and%20FLAG_LEVEL%20eq%202")
response <- GET(url)
resultTxt <- content(response, "text")
resultJson <- fromJSON(resultTxt, flatten = T)
yll_u70 <- resultJson$value %>% rename(FLAG_CAUSEGROUP = DIM_CAUSE_GROUP)

yll_u70_ <- yll_u70 %>%
  filter(DIM_COUNTRY_CODE == country_iso & FLAG_LEVEL==2 & DIM_YEAR_CODE==2021 &
           DIM_AGEGROUP_CODE %!in% c("Y70T74", "Y75T79", "Y80T84", "YGE_85", "TOTAL", "D0T27", "M1T11")) %>% 
  group_by(DIM_COUNTRY_CODE, DIM_YEAR_CODE, DIM_GHECAUSE_TITLE) %>% 
  summarise(yll_u70 = round(sum(VAL_YLL_COUNT_NUMERIC),0)) %>%
  ungroup() 

# yll_ %>% 
#   filter(DIM_YEAR_CODE==2021) %>% 
#   left_join(yll_u70_) %>% 
#   arrange(desc(yll)) %>% 
#   select(`Cause` = DIM_GHECAUSE_TITLE,
#          `Total` = yll,
#          `Under 70 years old` = yll_u70) %>% 
#   filter(row_number()<=15) %>% 
#   flextable() %>% 
#   width(., j =1, width=6, unit = "cm") %>%
#   width(., j =2:3, width=4, unit = "cm") %>%
#   align(., j=2:3, align = "center") %>% 
#   align(., j=2:3, align = "center", part = "header") %>% 
#   bold(., bold = TRUE, part = "header") %>% 
#   fontsize(., size = 9, part = "all") %>% 
#   bg(., i = NULL, j = NULL, "#e3ebf4", part = "header") %>% 
#   border_remove(.) %>% 
#   hline_bottom(., border = fp_border(color = "grey", width = 1), part = "body") %>%
#   hline_top(., border = fp_border(color = "grey", width = 1), part = "all")

# yll_ %>%
#   filter(DIM_YEAR_CODE==2021) %>%
#   mutate(age = "Total") %>% 
#   select(DIM_GHECAUSE_TITLE, age, yll) %>% 
#   rbind(yll_u70_ %>% select(DIM_GHECAUSE_TITLE, yll = yll_u70) %>% mutate(age = "Under 70 years old")) %>% 
#   arrange(desc(yll))

rank_order_yll <- yll_ %>% arrange(rank_yll) %>% pull(DIM_GHECAUSE_TITLE)

yll_ %>%
  filter(DIM_YEAR_CODE==2021) %>%
  left_join(yll_u70_) %>%
  arrange(desc(yll)) %>% 
  select(DIM_GHECAUSE_TITLE, yll, rank_yll, yll_u70) %>% 
  pivot_longer(cols = c(yll, yll_u70)) %>% 
  mutate(DIM_GHECAUSE_TITLE = factor(DIM_GHECAUSE_TITLE, levels = rank_order_yll)) %>% 
  filter(rank_yll<=10) %>% 
  mutate(age = ifelse(name=="yll_u70", "Under 70 years old", "Total")) %>% 
  ggplot(aes(x=DIM_GHECAUSE_TITLE, y = value, fill = age))+
  geom_col(position = position_dodge(), width = 0.6)+
  scale_fill_manual(values = c("Under 70 years old" ="#a5c6ec",
                               "Total" = "#009ADE"))+
  scale_x_discrete(limits=rev) +
  scale_y_continuous(labels = scales::number_format(big.mark = " "))+
  theme_void() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 9),
    panel.grid.major.x = element_line(color = "grey"),
    axis.text=element_text(size=9),
    axis.text.x = element_text(angle = 0),
    axis.text.y = element_text(hjust = 0)
  )+
  coord_flip()

```
\tiny Data source: WHO Global Health Estimates 2024
\newpage
```{r sdg prep, include=F, message=F, warning=F}
sdg_ref <- read.csv("C:\\Users\\samra\\Documents\\DNA\\country page pdf (datadot)\\sdg_ref_final.csv")
dec_ref <- readxl::read_xlsx("C:\\Users\\samra\\Documents\\DNA\\country page pdf (datadot)\\Rounding_WHS2024.xlsx")

getind <- function(){
  ind <- GET(url = "https://ghoapi.azureedge.net/api/Indicator")
  ind2 = fromJSON(rawToChar(ind$content))
  ind3 <- ind2[2]$value
  return(ind3)
}
ind_func <- getind()

url <- paste0("https://xmart-api-public.who.int/DATA_/RELAY_MAY2023?$filter=DIM_GEO_CODE_M49%20eq%20'", m49code, "'")
response <- GET(url)
resultTxt <- content(response, "text")
resultJson <- fromJSON(resultTxt, flatten = T)
sdgtest_ <- resultJson$value %>% 
  select(- c("_RecordID", "Sys_RowTitle", "Sys_Version", "Sys_VersionID", "Sys_OriginCode", "Sys_LoadBy", "Sys_CommitDateUtc", 
             "Sys_FirstLoadBy", "Sys_FirstCommitDateUtc", "Sys_ID", "Sys_BatchID", "Sys_FirstBatchID", "Sys_PK", "OBSERVATION_STATUS")) %>% 
  filter(DIM_TIME_TYPE != "YEAR_RANGE") %>% 
  filter(!is.na(VALUE_NUMERIC)) %>% 
  mutate(DIM_TIME = as.numeric(DIM_TIME)) %>% 
  filter(DIM_TIME>=2000 & DIM_TIME<2025) %>% 
  left_join(ind_func %>% select(-c(Language)) %>% rename(IND_CODE = IndicatorCode)) %>% 
  arrange(IndicatorName, DIM_TIME) %>% 
  left_join(dec_ref %>% select(IND_CODE = DimensionMemberCode, decimals) %>% mutate(decimals = as.numeric(decimals))) %>% 
  mutate(VALUE_NUMERIC = round(VALUE_NUMERIC, decimals))

# sdg_ref <- sdgtest_ %>% 
#   distinct(IND_CODE, IND_UUID, IndicatorName, DIM_MEMBER_1_CODE, DIM_MEMBER_2_CODE)
# write.csv(sdg_ref, "sdg_ref.csv")

```
\normalsize Indicator progress, 2000-2023
\vspace{-0.5cm}
```{r health status, fig.height=10, fig.width = 8, fig.align='center', message=F, warning=F}

if(who_region =="AFR"){
  ind_hs <- c("MALARIA_EST_INCIDENCE", "MDG_0000000020", "SDGHIV", "MDG_0000000026", "MDG_0000000007", "WSH_SANITATION_SAFELY_MANAGED")
}else if(who_region=="AMR"){
  ind_hs <- c("NCD_HYP_PREVALENCE_A", "VIOLENCE_HOMICIDERATE", "NCD_BMI_PLUS2C","SA_0000001688", "WSH_DOMESTIC_WASTE_SAFELY_TREATED")
}else if(who_region=="EMR"){
  ind_hs <- c("NCD_HYP_PREVALENCE_A", "SDGPM25", "SDGHEPHBSAGPRV", "SDGPOISON", "NCD_BMI_PLUS2C")
}else if(who_region=="EUR"){
  ind_hs <- c("SA_0000001688", "M_Est_tob_curr_std", "NCD_HYP_PREVALENCE_A", "SDGPOISON", "SDGSUICIDE")
}else if(who_region=="SEAR"){
  ind_hs <- c("SDGPM25", "MDG_0000000020", "MDG_0000000026", "NUTRITION_WH_2", "WHOSIS_000003", "WSH_WATER_SAFELY_MANAGED")
}else if(who_region=="WPR"){
  ind_hs <- c("MDG_0000000020", "NCD_BMI_30A", "SDGAIRBODA", "M_Est_tob_curr_std", "SDGSUICIDE")
}
  
sdg_hstat <- sdg_ref %>% 
  # filter(IND_CODE %in% c("MALARIA_EST_INCIDENCE", "MDG_0000000020", "SDGHIV", "WHS4_100", "M_Est_tob_curr_std", "NCD_HYP_PREVALENCE_A")) %>% 
  filter(IND_CODE %in% ind_hs) %>% 
  select(IND_UUID, DIM_MEMBER_1_CODE, DIM_MEMBER_2_CODE, increase_good, sort) %>% 
  left_join(sdgtest_)

hstat <- rbind(
  sdg_hstat %>% group_by(IND_UUID) %>% top_n(1, DIM_TIME) %>% mutate(year = "new"),
  sdg_hstat %>% group_by(IND_UUID) %>% top_n(-1, DIM_TIME)%>% mutate(year = "old")
) %>% 
  select(IND_UUID, IND_CODE, year, VALUE_NUMERIC, sort) %>% 
  pivot_wider(values_from = VALUE_NUMERIC, names_from = year) %>% 
  mutate(perc = ifelse(old!=0 & new!=0, ((new-old)/old)*100, 0)) %>% 
  mutate(flagz_s = ifelse(old==0 & new!=0, 1, 0),
         flagz_e = ifelse(old!=0 & new==0, 1, 0)) %>% 
  mutate(perc_abs = abs(perc)) %>% 
  arrange(sort) %>% 
  ungroup() %>% 
  mutate(row = row_number())

ind_hstat <- hstat %>% distinct(IND_UUID) %>%  pull(IND_UUID)

hstat_plots <- list()

for (i in ind_hstat){
  
  message(i)
  name <- sdg_ref %>% filter(IND_UUID==i) %>% pull(website_name) ; message(name)
  plot_num <- hstat %>% filter(IND_UUID==i) %>% distinct(row) %>% pull(row); message(plot_num)
  
  ind_code <- hstat %>% filter(IND_UUID==i) %>% pull(IND_CODE)
  dec_num <- dec_ref %>% filter(DimensionMemberCode==ind_code) %>% pull(decimals); dec_num <- as.numeric(dec_num)
  dec_num2 <- case_when(
    dec_num==0 ~ 1,
    dec_num==1 ~ 0.1,
    dec_num==2 ~ 0.01
    )
  
  data_i <- sdg_hstat %>% filter(IND_UUID==i) %>% 
    mutate(value_label = as.character(scales::number(VALUE_NUMERIC, accuracy = dec_num2,  big.mark = " ")))
  
  min_year <- min(data_i %>% pull(DIM_TIME))
  max_year <- max(data_i %>% pull(DIM_TIME))
  mid_year <- round((max_year+min_year)/2, digits = 0)
  
  mid_value <- (max(data_i %>% pull(VALUE_NUMERIC))+min(data_i %>% pull(VALUE_NUMERIC)))/2


  plot <- data_i %>% 
    ggplot(aes(x=DIM_TIME, y = VALUE_NUMERIC)) +
    labs(title = str_wrap(name, 50))+
    geom_line(color = "#009ADE", size = 1) +
    geom_point(color = "#009ADE", size = 1.3) +
    geom_point(data = data_i %>% filter(DIM_TIME %in% c(min_year, max_year)),
               aes(x=DIM_TIME, y = VALUE_NUMERIC),
               color = "#009ADE", size = 2) +
    geom_text_repel(data = data_i %>% filter(DIM_TIME %in% c(min_year, max_year)),
      aes(x=DIM_TIME, y = VALUE_NUMERIC, label = value_label)) +
    scale_x_continuous(limits = c(2000, 2025), breaks = c(min_year, max_year), n.breaks = 3)+
    scale_y_continuous(labels = scales::number_format(accuracy = dec_num2, big.mark = " "))+
    theme_classic()+
    theme(
    plot.title = element_text(size = 10),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    panel.border = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_line(color = "darkgrey",linetype = 3),
    panel.grid.major = element_line(color = "darkgrey",size = 0.5,linetype = 3)
  )
  
  hstat_plots[[plot_num]] <- plot
}


wrap_plots(hstat_plots[1:length(hstat_plots)], ncol = 2, axes = "keep")

```
\tiny Data source: WHO, The Global Health Observatory
\newpage
\normalsize Triple Billions - Number of people affected (millions)
```{r tripbil, fig.height=3.5, fig.width = 8, fig.align='center'}

url <- paste0("https://xmart-api-public.who.int/DATA_/RELAY_3B_DATA?$filter=DIM_GEO_CODE_M49%20eq%20'", m49code, "'")
response <- GET(url)
resultTxt <- content(response, "text")
resultJson <- fromJSON(resultTxt, flatten = T)
tripb <- resultJson$value 

tripb_ <- tripb %>% 
  group_by(DIM_GEO_CODE_M49, TRIPLE_BILLION, DIM_TIME) %>% 
  summarise(contrib = sum(COUNT_N)) %>% 
  ungroup() %>% 
  mutate(year = as.numeric(DIM_TIME)) %>% 
  mutate(tb = case_when(
    TRIPLE_BILLION=="HPOP" ~ "Healthier Populations",
    TRIPLE_BILLION=="HEP" ~ "Health Emergencies Protection",
    TRIPLE_BILLION=="UHC" ~ "Universal Health Coverage"
  )) %>% 
  mutate(contrib_m = round(contrib/1000000, 1))

tripb_  %>% 
  filter(year<=2025) %>% 
  ggplot() +
  geom_line(aes(x=year, y = contrib_m, color = tb), show.legend = F, size = 1)+
  geom_point(data = tripb_ %>% filter(year==2025),
             aes(x=year, y = contrib_m, color = tb), show.legend = F, size = 2)+
  geom_text_repel(data = tripb_ %>% filter(year==2025),
            aes(x = 2026, y = contrib_m, label = paste0(contrib_m, "m \n(", str_wrap(tb, 30), ")"), color = tb), 
            show.legend = F, size = 3, direction = "y",
            segment.color = 'transparent') +
  scale_x_continuous(breaks = c(2018,2019, 2020, 2021, 2022, 2023, 2024, 2025), limits = c(2018, 2027)) +
  scale_y_continuous(n.breaks = 5, labels = scales::label_number(suffix = "m")) +
  scale_color_manual(values = c("Healthier Populations" = "#009ADE",
                                "Health Emergencies Protection" = "#A6228C",
                                "Universal Health Coverage" = "#5B2C86"))+
  theme_classic() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    panel.border = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_line(color = "darkgrey",linetype = 3),
    panel.grid.major = element_line(color = "darkgrey",size = 0.5,linetype = 3),
    legend.title = element_blank(),
    legend.position = "top"
  )

uhc_contrib <- tripb_ %>% filter(DIM_TIME=="2025" & TRIPLE_BILLION=="UHC") %>% pull(contrib_m)
hep_contrib <- tripb_ %>% filter(DIM_TIME=="2025" & TRIPLE_BILLION=="HEP") %>% pull(contrib_m)
hpop_contrib <- tripb_ %>% filter(DIM_TIME=="2025" & TRIPLE_BILLION=="HPOP") %>% pull(contrib_m)
```
\small In `r country_`, by 2025 the number of additional people projected to be: enjoying better health and wellbeing is `r hpop_contrib` million; covered by essential services and not experiencing financial hardship is `r uhc_contrib` million; and protected from health emergencies is `r hep_contrib` million.
\rule{\textwidth}{0.6pt}\color{whodarkblue}
```{r uhccontrib, fig.align='center', fig.height=3.9, fig.width=7.5, include=FALSE}
# \normalsize Universal health coverage - individual indicator contributions (millions)

all_plt_dat2 <- all_plt_dat %>% 
      filter(
        country == country_,
        scenario == "baseline",
        ind %in% inds_all,
        year >= config_all$required_start_year,
        year <= 2030,
        year > 2017
      ) %>%
      left_join(ind_labels, by = c("ind", "billion")) %>%
      mutate(name = case_when(name == "Child healthcare" ~ "Child health care", .default = name))

ind_order <- all_plt_dat2 %>%
  filter(year == 2025 & billion == "uhc" & ind %!in% c("uhc_billion")) %>%
  mutate(
      pct_contribution = abs(contribution_mln) / sum(abs(contribution_mln)) * 100,
      ind_label = case_when(
        pct_contribution > 3 & contribution_mln>0 ~ name,
        pct_contribution > 3 & contribution_mln < 0 ~ name,
        .default = NA
      )
    ) %>% 
  arrange(abs(contribution)) 

uhc_ind_order <- ind_order %>%
  filter(!is.na(ind_label)) %>% 
  pull(ind_label)

ind_sent <- ind_order %>%
  filter(is.na(ind_label)) %>% 
  pull(name) %>% 
  paste(., collapse = ', ')

uhc_decomp <- all_plt_dat2 %>%
  filter(year<=2025 & billion == "uhc") %>% 
  mutate(ind_ = ifelse(name %in% c(uhc_ind_order, "UHC Billion"), name, "Other")) %>% 
  group_by(ind_, year) %>% 
  summarise(contribution_mln = sum(contribution_mln)) %>% 
  ungroup() %>%  
  mutate(ind_ = factor(ind_, levels = c(uhc_ind_order, "Other", "UHC Billion"))) %>% 
  rename(ind = ind_)

plt_dat = uhc_decomp
label_years = config_all$years_label_plot %>% c(2019)
size_text = 2.5
plt_billion = "uhc"
plt_net_ind = "UHC Billion"

net_dat <- uhc_decomp %>%
  filter(ind == "UHC Billion") %>%
  mutate(
    point_label = case_when(
      year %in% label_years ~ contribution_mln %>% round(digits = 0),
      .default = NA_real_
      )
  ) %>%
  select(ind, year, contribution_mln, point_label)
  
calc_label_points <- function(dat) {
    dat %>%
      mutate(
        ind_idx = as.integer(ind),
        ind_label = ind
      ) %>%
      arrange(-ind_idx) %>%
      mutate(
        y_upper_bar = cumsum(contribution_mln),
        y_lower_bar = c(0, y_upper_bar[-n()]),
        y_midpoint = (y_upper_bar + y_lower_bar) / 2,
      ) %>%
      select(ind, ind_idx, ind_label, year, contribution_mln, y_upper_bar, y_lower_bar, y_midpoint)
  }



label_positive_points <- plt_dat %>%
  filter(year == 2025 & contribution_mln >= 0 & ind != plt_net_ind) %>%
  calc_label_points()
label_negative_points <- plt_dat %>%
  filter(year == 2025 & contribution_mln < 0 & ind != plt_net_ind) %>%
  calc_label_points()
label_points <- bind_rows(label_positive_points, label_negative_points) 
  
plt_dat_ <- plt_dat %>%
  filter(ind != plt_net_ind) %>%
  mutate(ind_label = if_else(year == 2025, ind, NA))
  
target_color <- "#F26829" 
below <- plt_dat_ %>% 
  filter(year==2025 & contribution_mln < 0 & ind!="Other") %>% 
  arrange(contribution_mln) %>% 
  distinct(ind) %>% pull()
num_colors <- length(below)
# Number of colors in the gradient
color_gradient <- colorRampPalette(c("#feece0", target_color))(num_colors)
names(color_gradient) <- below

target_color <- "#009ADE"; # "#014461";
above <- plt_dat_ %>% 
  filter(year==2025 & contribution_mln > 0 & ind!="Other") %>% 
  arrange(desc(contribution_mln)) %>% 
  distinct(ind) %>% pull()
num_colors <- length(above)
# Number of colors in the gradient
color_gradient2 <- colorRampPalette(c("#ddeff9", target_color))(num_colors)
names(color_gradient2) <- above

color_final <- c(color_gradient2, color_gradient, "Other" = "#d5d5d5")
nudgeing <- net_dat %>% filter(year==2019) %>% pull(contribution_mln)

plt_dat_ %>%
    ggplot(aes(x = year, y = contribution_mln)) +
    geom_col(aes(fill = ind), position = "stack", na.rm = F, show.legend = F) + #, color = "#c5c6c7"
    geom_line(data = net_dat, aes(y = contribution_mln, color = "Net Contribution"),  color = "black") +
    geom_point(data = net_dat, aes(y = contribution_mln, color = "Net Contribution"), color = "black") +
    geom_text_repel(
      data = label_points %>% filter(ind_label!="Population change effort"),
      mapping = aes(x = 2025.45, y = y_midpoint, label = str_wrap(ind_label, 25)), #, color = ind
      color = "black",
      # vjust aligns the segments vertically in the middle of each bar section
      # position = position_stack(vjust = 0.5),
      na.rm = T,
      # align all indicator labels at 20301 onwards and only shift vertically
      direction = "y",
      # nudge_y = 1,
      xlim = c(2026, Inf),
      # still label indicators that are close to equal
      max.overlaps = 100,
      # effectively skip drawing line to label
      # min.segment.length = 100,
      show.legend = FALSE,
      size = size_text,
      lineheight = 0.75
    ) +
  geom_text_repel(
    data = net_dat %>% filter(year==2019),
      aes(x = year, y = contribution_mln, label = "Net contribution"),
      show.legend = F, na.rm = T, size = 3, color = "black",
      nudge_y = nudgeing
    ) +
    geom_hline(yintercept = 0, linetype = "solid", alpha = 0.5) +
    scale_fill_manual(
      values = color_final
    ) +
    scale_x_continuous(breaks = 2018:2025, expand = c(0,0)) +
    scale_y_continuous(n.breaks = 10) +
    theme_classic() +
    # Allow labels to bleed past the canvas boundaries
    coord_cartesian(clip = 'off') +
    theme(
      panel.grid.major.y = element_line(color = "darkgrey"),
      axis.line.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.line.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.x=element_text(size=8),
      axis.title.x =element_blank(),
      axis.text.y=element_text(size=8),
      axis.title.y=element_blank(),
      plot.title = element_blank(),
      plot.margin = unit(c(5, 120, 5, 5), "pt"),
    )
```
\normalsize Universal health coverage tracer indicator progress, 2018–2030
\
\includegraphics[width=7cm]{legend_uhc.png}
```{r uhccircle, fig.align='center', fig.height=4, fig.width=7}

donut_data <- filtered_indicator_values %>% 
      filter(billion == "uhc") %>%
      filter(country == country_) %>%
      filter(year %in% c(2018, 2030)) %>% 
      mutate(year_ = case_when(year==2018 ~ "start_year", year==2030 ~ "goal_year")) %>% 
      select(aggregate_id, year_, indicator_name, gpw13_indicator_code, billion, value, sdg_2030_goal) %>% 
      pivot_wider(names_from = year_, values_from = value) %>% 
      mutate(start_year = round(start_year, 0),
             goal_year = round(goal_year, 0)) %>% 
      filter(gpw13_indicator_code != "uhc_sm" &
               gpw13_indicator_code != "asc") %>%
      mutate(indicator_name = case_when(indicator_name == "Child treatment" ~ "Child Health Care Seeking",
                                        indicator_name == "Health security" ~ "Preparedness",
                                        .default = indicator_name)) %>% 
      arrange(goal_year) %>% 
  filter(goal_year!="NaN")


ind <- donut_data$indicator_name

  donut_data$indicator_name <- factor(donut_data$indicator_name, levels = c(ind))

  donut_data$xmin <- 0:(nrow(donut_data)-1)
  donut_data$xmax <- donut_data$xmin + 1
  
  scale_max <- max(donut_data$start_year, donut_data$goal_year, 120)

test <- ggplot(data = donut_data) +
    geom_col(aes(x=indicator_name, y = goal_year), fill = "#009ADE", alpha = 0.5, width = 1)+
    geom_col(aes(x=indicator_name, y = start_year), fill = "#009ADE", alpha = 1, width = 0.6) +
    geom_rect(aes(xmin = xmin+0.5, xmax = xmax+0.5,
                  ymin = sdg_2030_goal, ymax = sdg_2030_goal+1), color = "#A6228C",  fill = "#A6228C", show.legend = FALSE) +
    # geom_vline(xintercept = c(0.5,1.5,2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5,10.5,11.5,12.5), 
    #          color = "darkgrey", linetype = "solid", size = 0.4) +
    geom_hline(yintercept = c(scale_max, 0),
             color = "darkgrey", linetype = "solid", size = 0.25)
  
if((who_region_name=="African Region" & country_ %!in% c("Lesotho")) | country_ %in% c("Somalia")){
  
  test <- test +
    geom_segment(aes(x = c(0.5,1.5,2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5,10.5,11.5, 12.5), 
                   xend  = c(0.5,1.5,2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5,10.5,11.5, 12.5), 
                   y = 0, 
                   yend = scale_max),
               color = "darkgrey", lwd = 0.25)
} else {
  
  test <- test +
    geom_segment(aes(x = c(0.5,1.5,2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5,10.5,11.5), 
                   xend  = c(0.5,1.5,2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5,10.5,11.5), 
                   y = 0, 
                   yend = scale_max),
               color = "darkgrey", lwd = 0.25)
  
}
  test +
    coord_polar() +
    theme_void() +
    theme(axis.text.x = element_text(size = 6.5, color = "black", margin = margin(t = 0, r = 0, b =0, l = 0)),
          axis.ticks.x = element_line(color = "black"),
          plot.margin = unit(c(0,0,0,0), "cm"))+
    scale_y_continuous(limits = c(-20, scale_max))+
   scale_x_discrete(labels = function(x) str_wrap(x, width = 10))
  
#   test <- donut_data %>%
#     mutate(n = row_number()) %>%
#     pivot_longer(cols = c(sdg_2030_goal, start_year, goal_year)) %>%
#     mutate(name2 = case_when(
#       name=="sdg_2030_goal" ~ "2030 target",
#       name=="start_year" ~ "2018 baseline",
#       name=="goal_year" ~ "2030 forecast"
#     ))
#   test2 <- ggplot() +
#     geom_point(data = test %>% filter(name != "sdg_2030_goal"), aes(x=n, y = value, color = name2), size = 8) +
#     geom_line(data = test %>% filter(name == "sdg_2030_goal"), aes(x=n, y = value, color = name2), size = 3) +
#     theme_classic()+
#     theme(legend.title = element_blank(),
#           legend.position = "top",
#           legend.text = element_text(size = 12, margin = margin(r = 40, t = 0, l = 0, b = 0, unit = "pt"))
#           )+
#     scale_color_manual(values = c("2030 target" = "#A6228C",
#                                   "2018 baseline" = "#009ADE",
#                                   "2030 forecast" = "#bdd6f2"))
# 
#   png("rmd/legend_uhc.png", height = 30, width = 450)
# leg <- get_legend(test2)
# as_ggplot(leg)
# dev.off()
```
\tiny Data Source: WHO Triple Billions
\newpage
\normalsize Healthier populations - individual indicator contributions (millions)
```{r hpopcontrib, fig.align='center', fig.height=3.9, fig.width=7.5}

ind_order <- all_plt_dat2 %>%
  filter(year == 2025 & billion == "hpop" & ind != "hpop_healthier") %>%
  mutate(
      pct_contribution = abs(contribution_mln) / sum(abs(contribution_mln)) * 100,
      ind_label = case_when(
        pct_contribution > 3 & contribution_mln>0 ~ name,
        pct_contribution > 3 & contribution_mln < 0 ~ name,
        .default = NA
      )
    ) %>% 
  arrange(abs(contribution)) 

hpop_ind_order <- ind_order %>%
  filter(!is.na(ind_label)) %>% 
  pull(ind_label)

ind_sent <- ind_order %>%
  filter(is.na(ind_label)) %>% 
  pull(name) %>% 
  paste(., collapse = ', ')

hpop_decomp <- all_plt_dat2 %>%
  filter(year<=2025 & billion == "hpop") %>% 
  mutate(ind_ = ifelse(name %in% c(hpop_ind_order, "Population healthier - HPOP Billion"), name, "Other")) %>% 
  group_by(ind_, year) %>% 
  summarise(contribution_mln = sum(contribution_mln)) %>% 
  ungroup() %>%  
  mutate(ind_ = factor(ind_, levels = c(hpop_ind_order, "Other", "Population healthier - HPOP Billion"))) %>% 
  rename(ind = ind_)


plt_billion = "hpop"
plt_net_ind = "Population healthier - HPOP Billion"
label_years = config_all$years_label_plot %>% c(2019)
size_text = 2.5
plt_dat = hpop_decomp

net_dat <- plt_dat %>%
  filter(ind == plt_net_ind) %>%
  mutate(
    point_label = case_when(
      year %in% label_years ~ contribution_mln %>% round(digits = 0),
      .default = NA_real_
      )
  ) %>%
  select(ind, year, contribution_mln, point_label)
  
label_positive_points <- plt_dat %>%
  filter(year == 2025 & contribution_mln >= 0 & ind != plt_net_ind) %>%
  calc_label_points()
label_negative_points <- plt_dat %>%
  filter(year == 2025 & contribution_mln < 0 & ind != plt_net_ind) %>%
  calc_label_points()
label_points <- bind_rows(label_positive_points, label_negative_points) 
  
plt_dat_ <- plt_dat %>%
  filter(ind != plt_net_ind) %>%
  mutate(ind_label = if_else(year == 2025, ind, NA))
  
target_color <- "#F26829" 
below <- plt_dat_ %>% 
  filter(year==2025 & contribution_mln < 0 & ind!="Other") %>% 
  arrange(contribution_mln) %>% 
  distinct(ind) %>% pull()
num_colors <- length(below)
# Number of colors in the gradient
color_gradient <- colorRampPalette(c("#feece0", target_color))(num_colors)
names(color_gradient) <- below

target_color <- "#009ADE"; # "#014461";
above <- plt_dat_ %>% 
  filter(year==2025 & contribution_mln > 0 & ind!="Other") %>% 
  arrange(desc(contribution_mln)) %>% 
  distinct(ind) %>% pull()
num_colors <- length(above)
# Number of colors in the gradient
color_gradient2 <- colorRampPalette(c("#ddeff9", target_color))(num_colors)
names(color_gradient2) <- above

color_final <- c(color_gradient2, color_gradient, "Other" = "#d5d5d5")
nudgeing <- net_dat %>% filter(year==2019) %>% pull(contribution_mln)

plt_dat_ %>%
  ggplot(aes(x = year, y = contribution_mln)) +
  geom_col(aes(fill = ind), position = "stack", na.rm = F, show.legend = F) + #, color = "#c5c6c7"
  geom_line(data = net_dat, aes(y = contribution_mln, color = "Net Contribution"),  color = "black") +
  geom_point(data = net_dat, aes(y = contribution_mln, color = "Net Contribution"), color = "black") +
    geom_text_repel(
      data = label_points,
      mapping = aes(x = 2025.45, y = y_midpoint, label = str_wrap(ind_label, 25)), #, color = ind
      color = "black",
      # vjust aligns the segments vertically in the middle of each bar section
      # position = position_stack(vjust = 0.5),
      na.rm = T,
      # align all indicator labels at 20301 onwards and only shift vertically
      direction = "y",
      # nudge_y = 1,
      xlim = c(2026, Inf),
      # still label indicators that are close to equal
      max.overlaps = 100,
      # effectively skip drawing line to label
      # min.segment.length = 100,
      show.legend = FALSE,
      size = size_text,
      lineheight = 0.75
    ) +
  geom_text_repel(
    data = net_dat %>% filter(year==2019),
      aes(x = year, y = contribution_mln, label = "Net contribution"),
      show.legend = F, na.rm = T, size = 3, color = "black",
      nudge_y = nudgeing
    ) +
    geom_hline(yintercept = 0, linetype = "solid", alpha = 0.5) +
    scale_fill_manual(
      values = color_final
    ) +
    scale_x_continuous(breaks = 2018:2025, expand = c(0,0)) +
    scale_y_continuous(n.breaks = 10) +
    theme_classic() +
    # Allow labels to bleed past the canvas boundaries
    coord_cartesian(clip = 'off') +
    theme(
      panel.grid.major.y = element_line(color = "darkgrey"),
      axis.line.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.line.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.x=element_text(size=8),
      axis.title.x =element_blank(),
      axis.text.y=element_text(size=8),
      axis.title.y=element_blank(),
      plot.title = element_blank(),
      plot.margin = unit(c(5, 120, 5, 5), "pt"),
    )
```
\small Indicators aggregated as "Other" include: `r ind_sent`.
\rule{\textwidth}{0.6pt}\color{whodarkblue}
\
\normalsize Health emergencies protection - individual indicator contributions (millions)
```{r hepcontrib, fig.align='center', fig.height=3.9, fig.width=7.5}

hep_ind_order <- all_plt_dat2 %>%
  filter(year == 2025 & billion == "hep" & ind %!in% c("hep_idx")) %>%
  arrange(abs(contribution)) %>%
  pull(name)

hep_decomp <- all_plt_dat2 %>%
  filter(year<=2025 & billion == "hep") %>% 
  mutate(ind = ifelse(is.na(ind), name, ind)) %>% 
  mutate(ind = factor(name, levels = c(hep_ind_order, "HEPI")))

plt_billion = "hep"
plt_net_ind = "HEPI"
# plt_inds_color_pal = hpop_ind_colors
label_years = config_all$years_label_plot %>% c(2019)
size_text = 3
plt_dat = hep_decomp
  
net_dat <- plt_dat %>%
  filter(ind == plt_net_ind) %>%
  mutate(
    point_label = case_when(
      year %in% label_years ~ contribution_mln %>% round(digits = 0),
      .default = NA_real_
      )
  ) %>%
  select(ind, year, contribution_mln, point_label)
  
calc_label_points <- function(dat) {
    dat %>%
      mutate(
        ind_idx = as.integer(ind),
        ind_label = ind
      ) %>%
      arrange(-ind_idx) %>%
      mutate(
        y_upper_bar = cumsum(contribution_mln),
        y_lower_bar = c(0, y_upper_bar[-n()]),
        y_midpoint = (y_upper_bar + y_lower_bar) / 2,
      ) %>%
      select(ind, ind_idx, ind_label, year, contribution_mln, y_upper_bar, y_lower_bar, y_midpoint)
  }
  
  label_positive_points <- plt_dat %>%
    filter(year == 2025 & contribution_mln >= 0 & ind != plt_net_ind) %>%
    calc_label_points()
  label_negative_points <- plt_dat %>%
    filter(year == 2025 & contribution_mln < 0 & ind != plt_net_ind) %>%
    calc_label_points()
  label_points <- bind_rows(label_positive_points, label_negative_points) %>%
    mutate(
      pct_contribution = abs(contribution_mln) / sum(abs(contribution_mln)) * 100,
      ind_label = case_when(
        plt_billion == "hep" ~ ind,
        pct_contribution > 1 ~ ind,
        .default = NA
      )
    )
  label_ind_in_legend <- label_points %>%
    filter(is.na(ind_label)) %>%
    arrange(ind) %>%
    pull(ind)
  
  plt_dat <- plt_dat %>%
    filter(ind != plt_net_ind) %>%
    mutate(ind_label = if_else(year == 2025, ind, NA)) %>%
    select(ind, year, ind_label, contribution_mln)
  
target_color <- "#F26829" 
below <- label_points %>% 
  filter(contribution_mln < 0) %>% 
  arrange(contribution_mln) %>% 
  distinct(ind) %>% pull()
num_colors <- length(below)
# Number of colors in the gradient
color_gradient <- colorRampPalette(c("#f9bda2", target_color))(3)
names(color_gradient) <- below

target_color <- "#009ADE"; # "#014461"
above <- label_points %>% 
  filter(contribution_mln > 0) %>% 
  arrange(desc(contribution_mln)) %>% 
  distinct(ind) %>% pull()
num_colors <- length(above)
# Number of colors in the gradient
color_gradient2 <- colorRampPalette(c("#a0ccdf", target_color))(3)
names(color_gradient2) <- above

color_final <- c(color_gradient2, color_gradient)
ind_sent <- paste(label_ind_in_legend, collapse = ', ')
nudgeing <- net_dat %>% filter(year==2019) %>% pull(contribution_mln)

plt_dat %>%
    ggplot(aes(x = year, y = contribution_mln)) +
    geom_col(aes(fill = ind), position = "stack", na.rm = F, show.legend = F) +
    geom_line(data = net_dat, aes(y = contribution_mln, color = "Net Contribution"),  color = "black") +
    geom_point(data = net_dat, aes(y = contribution_mln, color = "Net Contribution"), color = "black") +
    geom_text_repel(
      data = label_points,
      mapping = aes(x = 2025.45, y = y_midpoint, label = ind_label), #, color = ind
      color = "black",
      # vjust aligns the segments vertically in the middle of each bar section
      # position = position_stack(vjust = 0.5),
      na.rm = T,
      # align all indicator labels at 20301 onwards and only shift vertically
      direction = "y",
      # nudge_y = 1,
      xlim = c(2026, Inf),
      # still label indicators that are close to equal
      max.overlaps = 100,
      # effectively skip drawing line to label
      # min.segment.length = 100,
      show.legend = FALSE,
      size = size_text
    ) +
  geom_text_repel(
    data = net_dat %>% filter(year==2019),
      aes(x = year, y = contribution_mln, label = "Net contribution"),
      show.legend = F, na.rm = T, size = 3, color = "black",
      nudge_y = nudgeing
    ) +
    geom_hline(yintercept = 0, linetype = "solid", alpha = 0.5) +
    scale_fill_manual(
      values = color_final,
      breaks = label_ind_in_legend
    ) +
    scale_x_continuous(breaks = 2018:2025, expand = c(0,0)) +
    scale_y_continuous(n.breaks = 10) +
    theme_classic() +
    # Allow labels to bleed past the canvas boundaries
    coord_cartesian(clip = 'off') +
    theme(
      panel.grid.major.y = element_line(color = "darkgrey"),
      axis.line.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.line.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.x=element_text(size=8),
      axis.title.x =element_blank(),
      axis.text.y=element_text(size=8),
      axis.title.y=element_blank(),
      plot.title = element_blank(),
      plot.margin = unit(c(5, 120, 5, 5), "pt"),
    )
```
\tiny Data Source: WHO Triple Billions
